### 2016-08-18 transformation.tex
improve flow in section 4

In the following, we illustrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al. \cite{ecrts15nelissen}, which allows analyzing the worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity. 
Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} can be applied under the following conditions:
\begin{enumerate}
	\item the task set contains only one self-suspending task, 
	\item the self-suspending task is the lowest-priority task, 
	\item the scheduling policy is preemptive fixed-priority scheduling, and 
	\item all tasks have constrained deadlines (i.e., $D_i \leq T_i$ for all $\tau_i$).
\end{enumerate}

Even for the restrictive case above, the mixed-integer linear programming (MILP) approach by Nelissen et al. \cite{ecrts15nelissen} does not provide the exact worst-case response time. 
However, if we apply the characteristics of the worst-case release pattern provided in Lemma 2 by Nelissen et al. \cite{ecrts15nelissen},  the exact worst-case response time can be calculated by exploring all  release patterns that satisfy these conditions. Detailed discussions can be found in a recent report by Chen \cite{Chen2016b}. In fact, it has recently been shown that verifying the schedulability of such a task set is coNP-hard in the strong sense even in the considered simplified case~\cite{RTSS2016-suspension}. Therefore, analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense by following the same proof \cite{RTSS2016-suspension}.

For an arbitrary number of tasks $k \geq 2$, 
suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines (i.e., $D_i = T_i$ for all $\tau_i$). Further suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.

@@

However, deriving an exact bound $W_k^j$ for $j=1,2,\ldots,m_k-1$ for task $\tau_k$ is not easy: 
even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound of the worst-case response time requires exponential time complexity if $j \geq 2$. 
Deriving the exact $W_k^j$ can be achieved by exploring all  release patterns that satisfy the conditions listed in Lemma 2 by Nelissen et al. \cite{ecrts15nelissen}, with very high time complexity. 
%Furthermore, Nelissen et al. \cite{ecrts15nelissen} identified several misconceptions in prior analyses, and after correcting those misconceptions, observed that the problem of deriving the worst-case response time of a computation segment in pseudo-polynomial time seems to be very challenging indeed.

Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} did not study the period enforcer; rather, they considered unrestricted self-suspensions. However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations transfer to the period enforcement case.

To summarize, to analyze the period enforcer based on Theorem~5 in~\cite{Raj:suspension1991}, a procedure for transforming multi-segmented self-suspending tasks into sets of single-segment deferrable tasks is needed, but no such procedure is given in the original proposal~\cite{Raj:suspension1991}.
%
Based on the presented considerations, we conclude that filling in this missing step is non-trivial and observe that the closest known solution by Nelissen et al.~\cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task. It thus remains unclear how Theorem~5 in~\cite{Raj:suspension1991} can be used for schedulability analysis of sets of multi-segmented self-suspending tasks. 
While we did search for alternative analysis approaches that do not rely on Theorem 5, we did not find a simple or efficient schedulability test for the period enforcer without introducing substantial additional pessimism.  The problem remains open.

>>>>

In the following, we illustrate the inherent difficulty of the problem by focusing on a special case to which we can apply a recent result of Nelissen et al. \cite{ecrts15nelissen}, which allows analyzing the worst-case response time of multi-segmented self-suspending sporadic tasks, albeit with exponential time complexity. Specifically,  
Nelissen et al.'s worst-case response time analysis~\cite{ecrts15nelissen} can be applied under the following conditions:
\begin{enumerate}
	\item the task set contains only one self-suspending task, 
	\item the self-suspending task is the lowest-priority task, 
	\item the scheduling policy is preemptive fixed-priority scheduling, and 
	\item all tasks have constrained deadlines (i.e., $D_i \leq T_i$ for all $\tau_i$).
\end{enumerate}

As an aside, it is interesting to note that, even for the restrictive case above,  Nelissen et al.'s approach \cite{ecrts15nelissen}, which is based on \emph{mixed-integer linear programming} (MILP), provides only an upper bound on the worst-case response time, and not necessarily the \emph{exact} worst-case response time. However, at least conceptually, the exact worst-case response time \emph{can} be obtained based on their analysis by combining it with an exhaustive search. Specifically, Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen} provides a characterization of the worst-case release pattern that yields a maximal response time. Thus, by exploring all  release patterns that satisfy the conditions stated in Lemma 2 of Nelissen et al.~\cite{ecrts15nelissen}, an exact bound can be determined, with very high time complexity. A detailed discussion of this approach can be found in a recent report~\cite{Chen2016b}.

Let us now return to the discussion of the task-set transformation needed to apply Theorem 5 in~\cite{Raj:suspension1991}. For an arbitrary number of tasks $k \geq 2$, 
suppose that the system has $k-1$ regular sporadic tasks and only one segmented self-suspending task $\tau_k$, and that all tasks have implicit deadlines (i.e., $D_i = T_i$ for all $\tau_i$). Further suppose that task $\tau_k$ has $m_k$ segments with $m_k \geq 3$.

@@

However, deriving an exact bound $W_k^j$ for task $\tau_k$ and for $j=1,2,\ldots,m_k-1$ is not easy: 
even for the above ``simple'' case, Nelissen et al.'s MILP solution~\cite{ecrts15nelissen} for calculating a safe upper bound on the worst-case response time requires exponential time complexity if $j \geq 2$. In fact, it has recently been shown that the problem of verifying the schedulability of such a task set is coNP-hard in the strong sense~\cite{RTSS2016-suspension}. Therefore, it follows analogously that analyzing the exact worst-case response time of task $\tau_k$ is NP-hard in the strong sense~\cite{RTSS2016-suspension}.

Notably, Nelissen et al.~\cite{ecrts15nelissen} and Chen \cite{RTSS2016-suspension} do not consider the period enforcer; rather, their analyses pertain to unrestricted self-suspensions. However, given that the period enforcer has no effect on tasks that do not self-suspend~\cite{Raj:suspension1991}, and given that in the considered special case only the lowest-priority task self-suspends, we believe that these observations transfer to the period enforcement case.

To summarize, to analyze the period enforcer based on Theorem~5 in~\cite{Raj:suspension1991}, a procedure for transforming multi-segmented self-suspending tasks into sets of single-segment deferrable tasks is needed, but no such procedure is given in the original proposal~\cite{Raj:suspension1991}.
%
Based on the presented considerations, we conclude that filling in this missing step is non-trivial and observe that the closest known solution by Nelissen et al.~\cite{ecrts15nelissen} requires exponential time even in the greatly simplified special case of a single self-suspending task, and that the problem of verifying the schedulability of such a task set is in fact coNP-hard in the strong sense~\cite{RTSS2016-suspension}. It thus remains unclear how Theorem~5 in~\cite{Raj:suspension1991} can be used for schedulability analysis of sets of multi-segmented self-suspending tasks. 
While we did search for alternative analysis approaches that do not rely on Theorem 5, we did not find a simple or efficient schedulability test for the period enforcer without introducing substantial additional pessimism.  The problem remains open.
